# What's the smallest amount you can't make with 5 coins ?

My amazing, awesome wife often comes up with the little puzzles for our
amazing children, and this one seemed destined to be solved in R.  So, 
using up to 5 coins (1p, 2p, 5p, 10p, 20p and 50p) first she asked our 
kids whether they could make every value up to 50p, and then what the 
smallest value they couldn't make was.

Here's my R solution (which took about 5mins less than our daughter took 
to answer the first question)

```{r}
# What Amounts can't you make using up to 5 coins 1p to 50p
# 
# Author: Paul Hurley
###############################################################################
library(ggplot2)
library(plyr)

#Define our coins
coins<-as.factor(c(0,1,2,5,10,20,50))

#build a list of all the possibilities
possibilities<-expand.grid(coin1=coins, coin2=coins, 
coin3=coins, coin4=coins, coin5=coins)

#calculate the result
possibilities$total<-as.numeric(as.character(possibilities$coin1))+
		as.numeric(as.character(possibilities$coin2))+
		as.numeric(as.character(possibilities$coin3))+
		as.numeric(as.character(possibilities$coin4))+
		as.numeric(as.character(possibilities$coin5))

#define our target values
targets<-1:250

#what amounts aren't possible
targets[!targets %in% possibilities$total]
```

So, the smallest value we can't make is 
`r min(targets[!targets %in% possibilities$total])`

We can even produce a table of the number of ways to make each
 number, and a graph
```{r}
tableofpossibilities<-ddply(.data=possibilities, .(total), nrow)

ggplot(data=possibilities, aes(x=total))+geom_histogram(binwidth=1)
```

Then when I triumphantly told her, she asked, 'what about 4 coins ?'
```{r}
#How about 4 coins
#build a list of all the possibilities
fourpossibilities<-expand.grid(coin1=coins, coin2=coins, 
		coin3=coins, coin4=coins)

#calculate the result
fourpossibilities$total<-as.numeric(as.character(fourpossibilities$coin1))+
		as.numeric(as.character(fourpossibilities$coin2))+
		as.numeric(as.character(fourpossibilities$coin3))+
		as.numeric(as.character(fourpossibilities$coin4))

#what values can't be made ?
targets[!targets %in% fourpossibilities$total]
```
So, the answer is `r min(targets[!targets %in% fourpossibilities$total])`and a graph
```{r}

ggplot(data=fourpossibilities, aes(x=total))+geom_histogram(binwidth=1)
```